Paper

### 2019-08-20 Paper.Rmd
...

Based on HGD/IMC/LGD

```{r pathroc, echo=F, fig.height=4, fig.width=12 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

@@

### Per Patient ROC

```{r echo=F, fig.height=4, fig.width=12}

# Excluding the final endoscopy
ptendo = do.call(bind_rows, lapply(unique(pg.samp$Hospital.Research.ID), function(pt) {
  x = pg.samp %>% filter(Hospital.Research.ID == pt)  
  x %>% filter(Endoscopy.Year != max(x$Endoscopy.Year))
})) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise_at(vars(Prediction), funs(max, mean), na.rm=T)

@@

@@

#back = back %>% filter(!is.na(Prediction))

@@

# Tumour Supressors

```{r ts, eval=F}
colnames(ccgenes)

tsgs = ccgenes
nrow(tsgs)

tsgs = makeGRangesFromDataFrame(tsgs[,c(colnames(loc), 'Gene.Symbol')], keep.extra.columns = T)
tsgs = sort(tsgs)

length(tsgs)

# average 1 per 9MB
avgTSGdist = mean(elementMetadata(distanceToNearest(tsgs))$distance)

chh = arrange(chh, -hazards)
features = makeGRangesFromDataFrame(chh, keep.extra.columns = T)

distanceTSG = data.frame(matrix(nrow=0,ncol=3))
for (i in 1:length(features)) {
  f = features[i]
  hits = nearest(f, tsgs, ignore.strand=T, select='all')
  if (length(hits) > 0) {
    distanceTSG = rbind(distanceTSG, cbind.data.frame(
      'features'=i,
      'nearestTSG'=elementMetadata(tsgs[subjectHits(hits)])$Gene.Symbol,
      'distanceToTSG'=elementMetadata(distanceToNearest(f, tsgs, ignore.strand=T, select='all'))$distance)
    )
  }
}

features[unique(subset(distanceTSG, distanceToTSG == 0)$features)]

# In our dataset we should see 1 TSG/onc per 10MB
# But this is double the actual number we see
ncol(dysplasia.df)/round(avgTSGdist/5e6)

length(which(round((subset(distanceTSG, distanceToTSG > 0)$distanceToTSG/5e6)-0.5) == round((avgTSGdist/5e6)-0.5)))
```

@@

@@

ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)

@@

```{r echo=F, eval=T}
dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions'
info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'

@@

files = list.files(dir, 'Rdata', recursive = T, full.names = T)

pfile = paste0(dir, '/ds_preds.txt')
if (file.exists(pfile)) {
  ds.preds = read_tsv(pfile)
} else {
  ds.preds = do.call(bind_rows, lapply(files, function(f) {
    load(f, verbose=T)
    bind_cols(predictions(prr), adjustRisk(prr, 'mean', 'sample')$adj.predictions %>% dplyr::select(Probability, `Relative Risk`) %>% dplyr::rename_all(funs(paste0), ' Adjusted ') )
  }))
  write_tsv(ds.preds, pfile)
}

@@

m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk','Relative Risk Adjusted'), id.vars = c('Type', 'Status', 'Risk'))

pr = ggplot(m, aes(variable, value)) + 
  facet_grid(~Type, scales = 'free_x') + 
  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
  #scale_color_gradientn(colors = myPal,  name='') +
  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
  theme(axis.text.x = element_text(angle=45, hjust=1))

>>>>

Based on HGD/IMC/LGD
```{r pathroc, echo=F, fig.height=4, fig.width=12 }
#sum(table(subset(patient.info, Status == 'P')$Pathology)[c('IMC','HGD')])/sum(table(subset(patient.info, Status == 'P')$Pathology))

@@

### Per Patient ROC
```{r echo=F, fig.height=4, fig.width=12}
# Excluding the final endoscopy
ptendo = do.call(bind_rows, lapply(unique(pg.samp$Hospital.Research.ID), function(pt) {
  x = pg.samp %>% filter(Hospital.Research.ID == pt)  
  x %>% filter(Endoscopy.Year != max(x$Endoscopy.Year))
})) %>% group_by(Patient, Hospital.Research.ID, Status) %>% summarise_at(vars(Prediction), funs(max, mean), na.rm=T)

@@

ggplot(back, aes(Status, RR, group=Status)) + 
  facet_grid(~Pathology) + 
  #scale_fill_manual(values=RColorBrewer::brewer.pal(4,"Paired")[c(1,3)]) + 
  scale_color_gradientn(colours = myPal) +
  geom_boxplot(outlier.shape = NA, fill='#BDBDBD') + 
  geom_jitter(width=0.2, aes(color=Prediction)) + 
  plot.theme  + theme(legend.position = 'none')

@@

@@

@@

## Normals sWGS sequenced

These normal tissues came from some samples that had been mistakenly sequenced from a timepoint (path ID) that Barrett's samples were also sequenced.  These include:

```{r, echo=F}
normal.info = readxl::read_xlsx(patient.file, sheet='Normals')

normal.info %>% dplyr::select(`Patient ID`, `Path ID`, `Pathology notes`) %>% kable(caption='Normals') %>% kable_styling('striped', full_width = F)
```

These were excluded from the model building procedure, but were then predicted using the resulting model.  We would expect them to all be low risk and all but one are. All have a relative risk below 0 as well.

```{r}

normal.preds = read_tsv('~/Data/BarrettsProgressionRisk/Analysis/multipcf_plots_normals/normal_predictions.tsv', col_types = cols(
  Sample = col_character(),
  Probability = col_double(),
  `Relative Risk` = col_double(),
  Risk = col_character(),
  Patient = col_character(),
  Hospital.Research.ID = col_character(),
  Path.ID = col_character(),
  Status = col_character(),
  Endoscopy = col_double(),
  Pathology = col_logical(),
  Plate.Index = col_character(),
  SLX.ID = col_double(),
  Barretts.Cellularity = col_logical(),
  P53.Status = col_character(),
  Total.Reads = col_double(),
  Batch.Name = col_character(),
  Set = col_character(),
  Block = col_double()
))

ggplot(normal.preds, aes(Patient, Probability, color=Risk)) + ylim(0,1) + geom_point(size=2) +
  scale_color_manual(values = BarrettsProgressionRisk::riskColors() ) + plot.theme

@@

#ggsave(filename='plots/val_leg.png', plot=valLeg, width=4, height=2, units='in', dpi=300)

@@

```{r echo=F, eval=T}
dir = '~/Data/BarrettsProgressionRisk/Analysis/downsampled/predictions/0.9'
info = '~/Data/BarrettsProgressionRisk/QDNAseq/all_downsampled/downsampled_ids.xlsx'

@@

ds.preds = read_tsv(paste0(dir, '/downsampled_preds.tsv'), col_types = cols(
  Sample = col_character(),
  Probability = col_double(),
  `Relative Risk` = col_double(),
  Risk = col_character(),
  PatientID = col_character(),
  Status = col_character(),
  Tissue = col_character(),
  `Expected Risk` = col_character(),
  Grade_patient = col_character(),
  Grade_biopsy = col_character(),
  Study = col_character(),
  Type = col_character(),
  Endoscopy = col_date(format = "")
))

@@

m = reshape2::melt(ds.preds, measure.vars = c('Relative Risk'), id.vars = c('Type', 'Status', 'Risk'))

pr = ggplot(m, aes(variable, value)) + 
  facet_grid(~Type, scales = 'free_x') + 
  geom_boxplot(aes(group=variable, fill=Type), outlier.colour = NA, show.legend = F) + 
  geom_jitter(width=0.2, aes(shape=Status, color=Risk), size=2, show.legend = F) + 
  scale_color_manual(values =  BarrettsProgressionRisk::riskColors()) +
  #scale_color_gradientn(colors = myPal,  name='') +
  scale_fill_manual(values=rev(RColorBrewer::brewer.pal(5,'RdYlBu')[c(1,2,4,5)])) +
  labs(x='', y='Relative Risk', title='Predictions on downsampled patients') + plot.theme +
  theme(axis.text.x = element_text(angle=45, hjust=1))
